\emph{hp}-Finite Element Model of Poisson and Nernst-Planck System of Equations

### 2010-10-07 model.tex
Finished the results section

\section{Model}
\begin{figure}
  \begin{centering}
  \includegraphics[scale=0.4]{domain}
  \caption{\label{fig:domain} Calculation domain $\Omega\subset\mathbb{R}^2$
  	with boundaries $\partial\Omega_{1\ldots 4}\subset\partial\Omega$.}
  \end{centering}
\end{figure}

The model for solving NPN system is implemented in Hermes2D. In this work
a rectangular 2D domain $\Omega\subset\mathbb{R}^2$ with boundaries 
$\partial\Omega_{1\ldots 4}\subset\partial\Omega$  is considered (see Fig.~\ref{fig:domain}). 
As there is now flow considered in or out of the domain and there is no
constant concentration at the boundaries, zero Neumann boundary conditions are used 
for Eq.~\eqref{eq:nernst-planck}:
\begin{equation}
  -D \frac{\partial C}{\partial n} - z \mu F C \frac{\partial \phi} {\partial n} = 0.
  \label{eq:nernst-planck-boundary}
\end{equation}
As a positive voltage $V_{pos}$ is applied on $\Omega_1$ and $V_{neg}=0$ is applied
on $\Omega_3$, Dirichlet boundary conditions are used for Eq.~\eqref{eq:poisson} for
boundaries $\Omega_1$ and $\Omega_3$:
\begin{eqnarray}
  \phi_{\partial\Omega_1}&=&V_{pos},\\
  \phi_{\partial\Omega_3}&=&0,
  \label{eq:dirichlet}
\end{eqnarray}
and Neumann boundaries for $\Omega_2$ and $\Omega_4$:
\begin{equation}
  \frac{\partial \phi_{\Omega_2}}{\partial n}=\frac{\partial \phi_{\Omega_4}}{\partial n}=0.
\end{equation}

@@

\subsection{Implementation in Hermes2D}
To implement the system of equations Eq.~\eqref{eq:nernst-planck-weak-final}
and Eq.~\eqref{eq:poisson-weak-final}, the residuals and the Jacobian
matrix must be derived. For that, Crank-Nicolson time stepping was
used
\begin{equation}
  \frac{\partial C}{\partial t} \approx \frac{C^{n+1} - C^n}{\tau},
  \label{eq:cranic}
\end{equation}
where $\tau$ is a time step. For the variables $C^{n+1}$ and $\phi^{n+1}$ the
following notation will be used:
\begin{eqnarray}
  C^{n+1} &=& \sum_{k=1}^{N^C} y_k^{C} v_k^{C}, \label{eq:cnotation}j\\
  \phi^{n+1} &=& \sum_{k=1}^{N^{\phi}} y_k^{\phi} v_k^{\phi}\label{eq:phinotation},
\end{eqnarray}
where $v_k^C$ and $v_k^\phi$ are piecewise polynomial functions in $V$.
Considering the Crank-Nicolson time stepping and the notation~\eqref{eq:cnotation},
the time discretized Eq.~\eqref{eq:nernst-planck-weak-final} becomes
\begin{eqnarray}
  F_i^C\left(Y\right) & = & \int_{\Omega} \frac{C^{n+1}}{\tau}v_i^C d\mathbf{x} - 
  \int_{\Omega} \frac{C^{n}}{\tau}v_i^C d\mathbf{x}\nonumber\\
  &&+\frac 12 \left[D\int_{\Omega} \nabla C^{n+1} \cdot \nabla v_i^C d\mathbf{x}+ 
  	D\int_{\Omega} \nabla C^{n} \cdot \nabla v_i^C d\mathbf{x}\right]\nonumber\\
  &&+ \frac 12 \left[K\int_{\Omega}C^{n+1} \left(\nabla \phi^{n+1} \cdot \nabla v_i^C\right) d\mathbf{x}+
  K\int_{\Omega}C^{n} \left(\nabla \phi^{n} \cdot \nabla v_i^C\right) d\mathbf{x}\right]\label{eq:Fc},
\end{eqnarray}
and in the notation~\eqref{eq:phinotation}, Eq.~\eqref{eq:poisson-weak-final} becomes
\begin{equation}
  F_i^{\phi}\left(Y\right) = \int_{\Omega} \nabla \phi^{n+1} \cdot \nabla v_i^{\phi} d\mathbf{x} 
  - \int_{\Omega} LC^{n+1}v_i^{\phi} d\mathbf{x} + \int_{\Omega} LC_0 v_i^{\phi} d\mathbf{x}.
  \label{eq:Fphi}
\end{equation}
For the implementation the $2\times 2$ Jacobian matrix $DF/DY$ with elements corresponding to
\begin{equation}
  \frac{\partial F_i^C}{\partial y_j^C}, \ \frac{\partial F_i^C}{\partial y_j^{\phi}},\  
  \frac{\partial F_i^{\phi}}{\partial y_j^C}, \ \frac{\partial F_i^{\phi}}{\partial y_j^{\phi}},
\end{equation}
must be derived: 
\begin{eqnarray}
  \frac{\partial F_i^C}{\partial y_j^C} &=& 
  \int_{\Omega} \frac{1}{\tau} v_j^C v_i^C d\mathbf{x} + 
  \frac 12 D\int_{\Omega} \nabla v_j^C \cdot \nabla v_i^C d\mathbf{x}
  + \frac 12 K\int_{\Omega} v_j^C \left(\nabla \phi^{n+1} \cdot \nabla v_i^C\right) d\mathbf{x},\label{eq:dFcdyc}\\
  \frac{\partial F_i^C}{\partial y_j^{\phi}} &=&
  \frac 12 K \int_{\Omega} C^{n+1} \left(\nabla v_j^{\phi} \cdot \nabla v_i^C\right) d\mathbf{x},\\
  \frac{\partial F_i^{\phi}}{\partial y_j^C} &=&
  - \int_{\Omega} L v_j^C v_i^{\phi} d\mathbf{x},\\
  \frac{\partial F_i^{\phi}}{\partial y_j^{\phi}} &=&
  \int_{\Omega} \nabla v_j^{\phi} \cdot \nabla v_i^{\phi} d\mathbf{x}\label{eq:dFphidyphi}.
\end{eqnarray}
In Hermes2D Eq.~\eqref{eq:Fc} and \eqref{eq:Fphi} define the residuum $F$ and
Eq.~\eqref{eq:dFcdyc}---\eqref{eq:dFphidyphi} define the Jacobian matrix $J$.

@@

The defined Jacobian $J$ and residuum $F$ can be simply solved in Hermes2D by
using Newton's iteration. However, we were more interested in the automatic
adaptivity and the calculation differences with a single-mesh and multi-mesh.

In case of the single mesh, the same basis function base was used, i.e.
$v_i^\phi=v_i^C$. For multi mesh, the meshes and therefore the basis functions
were different. Regardless of the meshing, automatic adaptivity was applied
during each time step, till the error converged to the acceptable level.
In analogy to the most successful adaptive ODE solvers, Hermes2D uses a pair of 
approximations with different orders of accuracy to obtain this information: 
coarse mesh solution and fine mesh solution. The initial coarse mesh is read 
from the mesh file, and the initial fine mesh is created through its 
global refinement both in $h$ and $p$. For more
information, see~\cite{Hermes-project}. In the calculations, the coarse
mesh solution was not calculated at each time step, but the fine mesh
solution was projected to the coarse mesh by using orthogonal projection. This
yielded better convergence and also faster calculation.

Hermes2D supports 8 different refinement modes, namely,
3 isotropic and 5 anisotropic refinements. The isotropic refinements are
\emph{h}-isotropic (H\_ISO), \emph{p}-isotropic (P\_ISO), \emph{hp}-isotropic (HP\_ISO).
Anisotropic refinement modes are
\emph{h}-anisotropic (H\_ANISO),
\emph{hp}-anisotropic-\emph{h} (HP\_ANISO\_H), \emph{p}-anisotropic (P\_ANISO),
\emph{hp}-anisotropic-p (HP\_ANISO\_P), and \emph{hp}-anisotropic (HP\_ANISO).
Once the refinement mode is selected (by user), Hermes2D selects a particular
refinement scheme from several candiates based on the score. More information
can be found in~\cite{Hermes-project}.

>>>>

\section{Model}
\begin{figure}
  \begin{centering}
  \includegraphics[width=0.2\columnwidth]{domain}
  \caption{\label{fig:domain} Calculation domain $\Omega\subset\mathbb{R}^2$
  	with boundaries $\partial\Omega_{1\ldots 4}\subset\partial\Omega$.}
  \end{centering}
\end{figure}

The model for solving NPN system is implemented in Hermes.
A rectangular 2D domain $\Omega\subset\mathbb{R}^2$ with boundaries 
$\partial\Omega_{1\ldots 4}\subset\partial\Omega$  is considered (see Fig.~\ref{fig:domain}). 
As there is now flow considered in or out of the domain and there is no
constant concentration at the boundaries, zero Neumann boundary conditions are used 
for Eq.~\eqref{eq:nernst-planck}:
\begin{equation}
  -D \frac{\partial C}{\partial n} - z \mu F C \frac{\partial \phi} {\partial n} = 0.
  \label{eq:nernst-planck-boundary}
\end{equation}
As a positive voltage $V_{pos}$ is applied on $\Omega_1$ and $V_{neg}=0$ is applied
on $\Omega_3$, Dirichlet boundary conditions are used for Eq.~\eqref{eq:poisson} for
boundaries $\Omega_1$ and $\Omega_3$:
\begin{eqnarray}
  \phi_{\partial\Omega_1}&=&V_{pos},\\
  \phi_{\partial\Omega_3}&=&0,
  \label{eq:dirichlet}
\end{eqnarray}
and Neumann boundaries for $\Omega_2$ and $\Omega_4$:
\begin{equation}
  \frac{\partial \phi_{\Omega_2}}{\partial n}=\frac{\partial \phi_{\Omega_4}}{\partial n}=0.
\end{equation}

@@

\subsection{Implementation in Hermes}
To implement the system of equations Eq.~\eqref{eq:nernst-planck-weak-final}
and Eq.~\eqref{eq:poisson-weak-final}, the residuals and the Jacobian
matrix must be derived. For that, Crank-Nicolson time stepping was
used
\begin{equation}
  \frac{\partial C}{\partial t} \approx \frac{C^{n+1} - C^n}{\tau},
  \label{eq:cranic}
\end{equation}
where $\tau$ is a time step. For the variables $C^{n+1}$ and $\phi^{n+1}$ the
following notation will be used:
\begin{eqnarray}
  C^{n+1} &=& \sum_{k=1}^{N^C} y_k^{C} v_k^{C}, \label{eq:cnotation}j\\
  \phi^{n+1} &=& \sum_{k=1}^{N^{\phi}} y_k^{\phi} v_k^{\phi}\label{eq:phinotation},
\end{eqnarray}
where $v_k^C$ and $v_k^\phi$ are piecewise polynomial functions in $V$.
Considering the Crank-Nicolson time stepping and the notation~\eqref{eq:cnotation},
the time discretized Eq.~\eqref{eq:nernst-planck-weak-final} becomes
\begin{eqnarray}
  F_i^C\left(Y\right) & = & \int_{\Omega} \frac{C^{n+1}}{\tau}v_i^C d\mathbf{x} - 
  \int_{\Omega} \frac{C^{n}}{\tau}v_i^C d\mathbf{x}\nonumber\\
  &&+\frac 12 \left[D\int_{\Omega} \nabla C^{n+1} \cdot \nabla v_i^C d\mathbf{x}+ 
  	D\int_{\Omega} \nabla C^{n} \cdot \nabla v_i^C d\mathbf{x}\right]\nonumber\\
  &&+ \frac 12 \left[K\int_{\Omega}C^{n+1} \left(\nabla \phi^{n+1} \cdot \nabla v_i^C\right) d\mathbf{x}+
  K\int_{\Omega}C^{n} \left(\nabla \phi^{n} \cdot \nabla v_i^C\right) d\mathbf{x}\right]\label{eq:Fc},
\end{eqnarray}
and in the notation~\eqref{eq:phinotation}, Eq.~\eqref{eq:poisson-weak-final} becomes
\begin{equation}
  F_i^{\phi}\left(Y\right) = \int_{\Omega} \nabla \phi^{n+1} \cdot \nabla v_i^{\phi} d\mathbf{x} 
  - \int_{\Omega} LC^{n+1}v_i^{\phi} d\mathbf{x} + \int_{\Omega} LC_0 v_i^{\phi} d\mathbf{x}.
  \label{eq:Fphi}
\end{equation}
For the implementation the $2\times 2$ Jacobian matrix $DF/DY$ with elements corresponding to
\begin{equation}
  \frac{\partial F_i^C}{\partial y_j^C}, \ \frac{\partial F_i^C}{\partial y_j^{\phi}},\  
  \frac{\partial F_i^{\phi}}{\partial y_j^C}, \ \frac{\partial F_i^{\phi}}{\partial y_j^{\phi}},
\end{equation}
must be derived: 
\begin{eqnarray}
  \frac{\partial F_i^C}{\partial y_j^C} &=& 
  \int_{\Omega} \frac{1}{\tau} v_j^C v_i^C d\mathbf{x} + 
  \frac 12 D\int_{\Omega} \nabla v_j^C \cdot \nabla v_i^C d\mathbf{x}
  + \frac 12 K\int_{\Omega} v_j^C \left(\nabla \phi^{n+1} \cdot \nabla v_i^C\right) d\mathbf{x},\label{eq:dFcdyc}\\
  \frac{\partial F_i^C}{\partial y_j^{\phi}} &=&
  \frac 12 K \int_{\Omega} C^{n+1} \left(\nabla v_j^{\phi} \cdot \nabla v_i^C\right) d\mathbf{x},\\
  \frac{\partial F_i^{\phi}}{\partial y_j^C} &=&
  - \int_{\Omega} L v_j^C v_i^{\phi} d\mathbf{x},\\
  \frac{\partial F_i^{\phi}}{\partial y_j^{\phi}} &=&
  \int_{\Omega} \nabla v_j^{\phi} \cdot \nabla v_i^{\phi} d\mathbf{x}\label{eq:dFphidyphi}.
\end{eqnarray}
In Hermes Eq.~\eqref{eq:Fc} and \eqref{eq:Fphi} define the residuum $F$ and
Eq.~\eqref{eq:dFcdyc}---\eqref{eq:dFphidyphi} define the Jacobian matrix $J$.

@@

The defined Jacobian $J$ and residuum $F$ can be simply solved with Hermes by
using Newton's iteration. However, we were more interested in the automatic
adaptivity to control the error and problem size.

In a traditional low-order FEM, refining an element is not algorithmically complicated,
and so the most difficult part is to find out what elements should be refined. 
To do this, various techniques ranging from rigorous guaranteed a-posteriori 
error estimates to heuristic criteria such as residual error indicators, 
error indicators based on steep gradients, etc are employed. 
Unfortunately, none of these approaches are suitable for real-life 
multiphysics coupled problems or higher-order finite element methods: 
rigorous guaranteed error estimates only exist for very simple problems 
(such as linear elliptic PDE), and moreover only for low-order finite elements 
(such as piecewise linear approximations). These heuristic techniques 
are not employed in Hermes since they may fail 
in non-standard situations, and because they lack a 
transparent relation to the true approximation error.
In order to obtain fast, usable adaptivity, one has to resort to adaptive \emph{hp}-FEM.
Automatic adaptivity in the \emph{hp}-FEM is substantially different from adaptivity 
in low-order FEM, since every element can be refined in many different ways. 
Fig.~\ref{fig:refinements} shows several illustrative refinement candidates for a fourth-order element.
\begin{figure}
  \begin{centering}
  \includegraphics[width=0.5\columnwidth]{refinements}
  \caption{\label{fig:refinements} Refinement candidates for a fourth-order
  element.}
  \end{centering}
\end{figure}
The number of possible element refinements is implementation-dependent. 
In general it is very low in \emph{h} or \emph{p} adaptivity, 
but much higher in \emph{hp}-adaptivity, and it rises even more when 
anisotropic refinements are enabled.
Hermes supports 8 different refinement modes, namely,
3 isotropic and 5 anisotropic refinements. The isotropic refinements are
\emph{h}-isotropic (H\_ISO), \emph{p}-isotropic (P\_ISO), \emph{hp}-isotropic (HP\_ISO).
Anisotropic refinement modes are
\emph{h}-anisotropic (H\_ANISO),
\emph{hp}-anisotropic-\emph{h} (HP\_ANISO\_H), \emph{p}-anisotropic (P\_ANISO),
\emph{hp}-anisotropic-p (HP\_ANISO\_P), and \emph{hp}-anisotropic (HP\_ANISO).
Once the refinement mode is selected (by user), Hermes selects a particular
refinement scheme from several candidates based on the score. More information
can be found in~\cite{Hermes-project}.

In case of the single mesh, the same basis function base was used, i.e.
$v_i^\phi=v_i^C$. For multi mesh, the meshes and therefore the basis functions
were different. Regardless of the meshing, automatic adaptivity was applied
during each time step, till the error converged to the acceptable level.
In analogy to the most successful adaptive ODE solvers, Hermes uses a pair of 
approximations with different orders of accuracy to obtain this information: 
coarse mesh solution and fine mesh solution. The initial coarse mesh is read 
from the mesh file, and the initial fine mesh is created through its 
global refinement both in $h$ and $p$. For more
information, see~\cite{Hermes-project}. In the calculations, the coarse
mesh solution was not calculated at each time step, but the fine mesh
solution was projected to the coarse mesh by using orthogonal projection. This
yielded better convergence and also faster calculation.

### 2010-10-07 results.tex
Finished the results section

\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO adaptivity was used).}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
with adaptivity type HP\_ANISO is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time, some ionic migration has already taken place, i.e.
concentration gradient near the $\partial\Omega_1$ and
$\partial\Omega_3$ has formed. The automatic mesh refinements
at different time steps are clearly visible --- especially
near the top boundary, where the concentration gradient is
moving in time (as seen in Fig.~\ref{fig:comsol-conc-volt}).

@@

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  adaptivities. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} CPU times in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  adaptivities. (notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivities.}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} CPU times in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivities.}
  \end{centering}
\end{figure}
To narrow down the list of adaptivity modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. Here \emph{p}-adaptivity
modes are compared separately as they use more refined mesh.
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} show the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of CPU time
and problem size.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
Here we see that the anisotropic adaptivities result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO have a very largy problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means here that the number of degrees of freedom in time converges
to so that $N_{dof}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most adaptivity modes gave significantly smaller results,
e.g. $t<<500\ s$ and $N_{dof} << 500$.

@@

@@

So far we have seen that HP\_ANISO results in the smallest problem size, but requires
quite a lot CPU time. At the same time, HP\_ANISO\_P is the fastest, whereas
it does not result in that small problem. When it comes to a large domain 
or 3D modeling the problem
size becomes the most important factor. Therefore, we consider HP\_ANISO the most
suitable adaptivity type to the given problem. Thus some ways to optimize the time
factor will be considered.

The performance of HP\_ANISO\_P with refined mesh and half-refined
mesh, and performance of HP\_ANISO with unrefined, half-refined, and refined
meshes was calculated (see Fig.~\ref{fig:mesh}).
The results are shown in Fig.~\ref{fig:refined-cpu} and Fig.~\ref{fig:refined-dof}.

TODO(more explanation) It can be seen that some initial refinemts help to reduce the CPU time,
however, it increases the problem size.

>>>>

\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphi}
  \caption{\label{fig:cphi} Concentration $C$
  and voltage $\phi$ at two different time steps
  (HP\_ANISO adaptivity was used). This is a 2D solution shown
  as in 3D, where the height indicates the value.}
  \end{centering}
\end{figure}
Example of the solution at $t=0.7\ s$ and $t=3.0\ s$ 
with adaptivity type HP\_ANISO is shown
in Fig.~\ref{fig:cphi}. The time $t=0.7\ s$ was chosen because
by that time, some ionic migration has already taken place, i.e.
concentration gradient near the $\partial\Omega_1$ and
$\partial\Omega_3$ has formed. The automatic mesh refinements
at different time steps are clearly visible --- especially
near the top boundary, where the concentration gradient is
moving in time (as seen in Fig.~\ref{fig:comsol-conc-volt}).

@@

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_dof}
  \caption{\label{fig:isoanisodof} NDOFs in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  adaptivities. (Notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoaniso_cpu}
  \caption{\label{fig:isoanisocpu} CPU times in case 
  of multi-mesh solutions with H\_ISO, H\_ANISO,
  HP\_ISO, HP\_ANISO, and single-mesh solution with HP\_ANISO\_H
  adaptivities. (notice log Y scale)}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_dof}
  \caption{\label{fig:isoanisopdof} NDOFs in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivities.}
  \end{centering}
\end{figure}
\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{isoanisop_cpu}
  \caption{\label{fig:isoanisopcpu} CPU times in case 
  of multi-mesh solutions with P\_ISO, P\_ANISO, and
  HP\_ANISO\_P adaptivities.}
  \end{centering}
\end{figure}
To narrow down the list of adaptivity modes for the given problem, first the
isotropic and anisotropic adaptivities were compared. Here \emph{p}-adaptivity
modes are compared separately as they use more refined mesh.
Fig.~\ref{fig:isoanisodof} and Fig.~\ref{fig:isoanisocpu} show the comparison
of H\_ISO, H\_ANISO, HP\_ISO, HP\_ANISO, and HP\_ANISO\_H modes in terms of CPU time
and problem size.
Fig.~\ref{fig:isoanisopdof} and Fig.~\ref{fig:isoanisopcpu} show the similar
comparsion for the P\_ISO, P\_ANISO, and HP\_ANISO\_P modes.
Here we see that the anisotropic adaptivities result in a reasonable problem
size and the problems solve within a reasonable calculation time. It is interesting
to note that HP\_ANISO results in the smallest problem size. Also, in
\emph{p}-adaptivity group, HP\_ANISO\_P results in the smallest problem size
at each time step, whereas P\_ISO and P\_ANISO have a very largy problem size
during the first time steps of the solution.
Here the term ``reasonable problem size''
means that the number of degrees of freedom in time converges
to so that $N_{dof}<500$, and the term ``reasonable calculation time''
means that the calculation (step $\tau=0.01\ s$, physical
time $t_{end}=3.0\ s$) time $t$ on a givem system was $t<500\ s$.
Although these parameters are empirical, they serve as an upper limit, given
that the most adaptivity modes gave significantly smaller results,
e.g. $t<<500\ s$ and $N_{dof} << 500$.

@@

\subsection{Optimizations}

@@

So far we have seen that HP\_ANISO results in the smallest problem size, but requires
quite a lot CPU time. At the same time, HP\_ANISO\_P is the fastest, whereas
it does not result in that small problem. When it comes to a large domain 
or 3D modeling the problem
size becomes the most important factor. Therefore, we consider HP\_ANISO the most
suitable adaptivity type to the given problem. Thus some ways to optimize the time
factor will be considered. The desirable output would be HP\_ANISO problem size
close to HP\_ANISO\_P CPU time. One way to optimize the problem
is to choose somewhat more refined initial mesh. Other way one could think
to optimize the problem would be to change
the refinement frequency in a time dependent way. Recall that up to this point,
the initial mesh has been loaded in the beginning of each time step. True,
by employing the optimizations, we already must know something about the problem
and its solution beforehand. However, it could still be practial when solving a real problem
in a large domain.

The problem size and CPU time with HP\_ANISO
and HP\_ANISO\_P adaptivities on more refined initial mesh (see Fig.~\ref{fig:mesh}~(c))
compared to the coarse initial mesh (Fig.~\ref{fig:mesh}~(a)) and 
HP\_ANISO\_P solution is shown in Fig.~\ref{fig:refined-cpu} and Fig.~\ref{fig:refined-dof}.
It can be seen that by using initially more refined mesh, et problem solving time
can be reduced in case of HP\_ANISO, at the same time, the problem size increases.
In this situation, HP\_ANISO\_P and HP\_ANISO perform equally well.

\begin{figure}
  \begin{centering}
  \includegraphics[width=\columnwidth]{unreffreq_dof}
  \caption{\label{fig:unreffreq-dof} NDOFs at each time step for
  HP\_ANISO and HP\_ANISO\_P with mesh unrefinement at each time step and at over each
  	time step after first $0.5\ s$ physical solution time.}
  \end{centering}
\end{figure}
The next proposed optimization involved changing the unrefinement frequency.
It is known that the concentration gradient $\partial C$ changes the most in the initial phase
of the calculation, therefore, the unrefinement after each time step was performed
until $t=0.5\ s$ (physical time). After that, the
unrefinement was performed in $0.10\ s$ interval.
It appears that this type
of optimization does not result in a stable solution, i.e. the problem size 
does not remain steady, but starts to oscillate depending on the unrefinement
frequency. This is demonstrated in 
Fig.~\ref{fig:unreffreq-dof}. Therefore this kind of optimization did not
prove to be efficient.

\subsection{More general results}
\begin{figure}
  \begin{centering}
  \includegraphics[width=.75\columnwidth]{cphiorders}
  \caption{\label{fig:cphi-orders} Solutions $C$ and $\phi$
  and corresponding polynomial degrees of the elements at
  $t=0.7\ s$. HP\_ANISO adaptivity type was used. The height
  in the solution graphs indicates the value.}
  \end{centering}
\end{figure}

Based on the results, cation concentration and voltage was calculated
for different boundary conditions.
For instance, when voltage $\phi_{\Omega_1}=0.5\frac{x}{width_{\Omega_1}}+0.5$, the
concentration gradient $\nabla C$ and the voltage gradient $\nabla \phi$ are no
longer effectively 1D.
The calculated $C$ and $\phi$ in $\Omega$ and corresponding meshes and polynomial
degrees of the elements are shown in Fig.~\ref{fig:cphi-orders}.
HP\_ANISO adaptivity type was used. Notice that the solution
is different to the one in Fig.~\ref{fig:cphi} and the adapted mesh and the
polynomial degrees are also more complicated than in Fig.~\ref{fig:poly}.
It must be also noted that in case of non uniform boundary conditions which
results in 2D problem, refined initial mesh is more efficient to use.
